Community Embeddings for Facebook Friend Suggestions\\ \thanks{The reported study was partially supported by RFBR grant \textnumero 20-07-00958. The paper was prepared within the framework of the HSE University Project Group Competition 2020-2022.}

### 2020-10-12 paper-ffs.tex
add facebook to title
add HSE tranks
from ProfPPanfilov
natbib does not have a citetitle
citet cites last name of first author, or similar

\documentclass[conference]{IEEEtran}

\usepackage{cite}
\usepackage[numbers]{natbib}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{xcolor}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{multirow}
\usepackage{mathtools}
\usepackage{commath}
\usepackage{caption}
\usepackage{subcaption}
\usepackage[super]{nth}
\usepackage{url}

@@

\title{Community Embeddings for Friend Suggestions}

@@

\begin{figure}[htbp]
    \centering
    \includegraphics[width=0.48\textwidth]{images/fb/graph_d4.png}
    \caption{\citet{social_circles_fb} graph plotted with networkx's spring layout and node colors representing community membership obtained by ComE BGMM+VI with $D=4$ \cite{networkx}.}
    \label{fig:fb_graph}
\end{figure}

@@

The function $ComE(G, K, \gamma, \ell, \zeta, D, m, \alpha, \beta)$ is the ComE BGMM+VI community embedding algorithm. It returns the node embedding $\Phi$, context embedding $\Phi'$, community assignment $\Pi$, and community embedding $(\Psi,\Sigma)$. ComE BGMM+VI is based on the ComE algorithm presented by \citeauthor{ComE} in their paper \citet{ComE} as algorithm 1 on page 381 of the publication's in-proceeding \cite{ComE}. ComE BGMM+VI takes a Bayesian approach to community embedding by utilizing  a Bayesian Gaussian mixture model (BGMM) and variational inference (VI) for community embedding and detection instead of a non-Bayesian Gaussian mixture model (GMM) with expectation maximization (EM) \cite{ComE_BGMM_GH}. %TODO citation to the ComE BGMM paper.

@@

The small \citet{KarateClubDS} dataset will be used to visually underline the motivation for using node embedding and community memberships generated by ComE BGMM+VI for generating friend suggestions for person nodes on the karate club graph \cite{KarateClubDS}.

\begin{figure}[htbp]
    \centering
    \includegraphics[width=0.4\textwidth]{images/karateclub/spring.png}
    \caption{\citet{KarateClubDS} plotted with networkx's spring layout \cite{KarateClubDS, networkx}.}
    \label{fig:karateclub_spring}
\end{figure}

Fig. \ref{fig:karateclub_spring} shows the 34 nodes and 78 edges of the \citet{KarateClubDS} dataset plotted as a graph with the Python network package networkx's spring layout.\cite{networkx} Spring layout simulates a force-directed representation of the network, by treating edges as springs pulling nodes together and treating nodes as repelling objects, sometimes called an anti-gravity force.\cite{networkx} The node classification, represented by node color, shown in Fig. \ref{fig:karateclub_spring} was obtained by running ComE with BGMM and VI with the hyperparameters presented in Table \ref{tab:params_karateclub}.

@@

As can be seen in Fig. \ref{fig:karateclub_spring}, the two communities in the \citet{KarateClubDS} dataset are correctly identified, despite the number of communities being initialized with $K=5$. Thanks to the Bayesian approach to community modeling and optimization through variational inference, three unused communities are dropped, which leaves the prominent two communities present in the dataset. The node classification obtained reflects the node classification published in the \citeyear{ComE} original ComE paper by \citeauthor{ComE}.\cite{ComE}

@@

\citet{KarateClubDS} is split into training and testing parts. For evaluation, the leave-one-out (LOO) method is used, where for each testing user one edge is left out, then if the left out edge is in the user's top-$N$ recommendations, a hit is counted, otherwise, a miss is recorded.

@@

\begin{figure}[htbp]
    \centering
    \includegraphics[width=1\linewidth]{images/karateclub/embeddings_5.png}
    \caption{\citet{KarateClubDS} training dataset node and community embedding and node classification.}
    \label{fig:karateclub_emb}
\end{figure}

@@

\begin{figure}[htbp]
    \centering
    \includegraphics[width=.4\textwidth]{images/karateclub/hit_rates.png}
    \caption{Hit-rates from top-$N$ recommendations on \citet{KarateClubDS}.}
    \label{fig:karateclub_hitrates}
\end{figure}

Fig. \ref{fig:karateclub_hitrates} shows promising results for using ComE to generate top-$N$ friend suggestions on the small \citet{KarateClubDS}. The achieved hit-rate starts at 33\% for $N=1$, reaches 67\% at $N=3$, and maxes out at 100\% for $N=7$; therefore, when presenting three friend suggestions to one test user, at least in $66\%$ of cases a friend suggestion would be of interest.

@@

The recommender system based on community embeddings with the Bayesian Gaussian mixture model and variational inference is quantitatively evaluated. Friend suggestions are generated for the \citet{social_circles_fb} dataset \cite{social_circles_fb, social_circles_in_ego_networks}.

The method of generating friend suggestions using ComE BGMM+VI embeddings and the smallest-distance approach is applied to the \citet{social_circles_fb} graph. Both Algorithm \ref{alg:friend_sug} and Algorithm \ref{alg:friend_sug_com} are evaluated. The top-$N$ recommendations are evaluated by the hit-rate metric.

@@

The \citet{social_circles_fb} dataset was sourced from the profile and network data of ten ego-networks (egonets) through a custom Facebook app by \citeauthor{social_circles_in_ego_networks} as part of their \citeyear{social_circles_in_ego_networks} article \citet{social_circles_in_ego_networks}.\cite{social_circles_in_ego_networks}

@@

The \citet{social_circles_fb} graph is split into training and testing parts. For evaluation, the leave-one-out (LOO) method is used, where for each testing user one edge is left out, then if the left out edge is in the user's top-$N$ recommendations, a hit is counted, otherwise, a miss is recorded.
Users with only one connection are removed from the dataset since no network information is added by them.

$1169$ of the 4039 users of the \citet{social_circles_fb} graph have at least $50$ connections and are used for testing friend suggestions. The threshold of $50$ friends for considering users for testing is set to make sure enough data is available to make sensable suggestions for testing users.

@@

Community embeddings for the \citet{social_circles_fb} graph are generated by running ComE BGMM+VI with the hyperparameters listed in Table \ref{tab:params_fb}.

\begin{table}
    \centering
    \caption{Hyperparameters used for the \citet{social_circles_fb} dataset.}
    \label{tab:params_fb}
    \begin{tabular}{ l | c | r }
        parameter                    & notation & value       \\
        \hline
        \hline
        number\_walks                & $\gamma$ & $10$        \\
        \hline
        walk\_length                 & $\ell$   & $80$        \\
        \hline
        representation\_size         & $D$      & $2, 4, 128$ \\
        \hline
        num\_workers                 & \empty   & $10$        \\
        \hline
        num\_iter                    & \empty   & $3$         \\
        \hline
        com\_n\_init                 & \empty   & $10$        \\
        \hline
        reg\_covar                   & \empty   & $0.00001$   \\
        \hline
        batch\_size                  & \empty   & $50$        \\
        \hline
        window\_size                 & $\zeta$  & $10$        \\
        \hline
        negative                     & $m$      & $5$         \\
        \hline
        lr                           & \empty   & $0.025$     \\
        \hline
        alpha                        & $\alpha$ & $0.1$       \\
        \hline
        beta                         & $\beta$  & $0.1$       \\
        \hline
        down\_sampling               & \empty   & $0.0$       \\
        \hline
        communities                  & $K$      & $10$        \\
        \hline
        weight\_concentration\_prior & $\Gamma$ & $10^{-5}$   \\
    \end{tabular}
\end{table}

As is shown in Table \ref{tab:params_fb}, the embedding's representation size $D$ is varied along three values. A representation size $D \in [2,4,128]$ is chosen to determine the effect of the embedding's dimensionality on friend suggestions. It is hypothesized, that two dimensions are not sufficient to capture the structure of the \citet{social_circles_fb} graph, due to its size, therefore higher dimensional embeddings are evaluated as well.

@@

Fig. \ref{fig:fb_emb_anim_d2} shows node and community embeddings with representation size $D=2$. The limiting two-dimensional space does not allow ComE to represent the \citet{social_circles_fb} graph's structure sufficiently to utilize the $K=10$ components of the Gaussian mixture model. Only three components are used after the last of three iterations.

@@

    \caption{Hit-rates of \citet{social_circles_fb} graph's top-$N$ friend suggestions on $N$-axis for $D \in [2,4,128]$. Hit-rate lines for Algorithm \ref{alg:friend_sug} and Algorithm \ref{alg:friend_sug_com} overlap for $D=2$ and $D=4$.}
    \label{fig:fb_hit_rates}
\end{figure}

@@

\begin{figure}[htbp]
    \centering
    \caption{Hit-rates of \citet{social_circles_fb} graph's top-$N$ friend suggestions for algorithms \ref{alg:friend_sug} and \ref{alg:friend_sug_com}, $N \in [5, 10, 20]$, and $D \in [2, 4, 128]$.}
    \label{tab:fb_hit_rates}
    \begin{tabular}{ l | l | c | c | c }
        \hline
        \empty                                              & \empty  & top-$5$         & top-$10$        & top-$20$        \\
        \hline
        \hline
        \multirow{3}{*}{Algorithm \ref{alg:friend_sug}}     & $D=2$   & 8.5\%           & 14.2\%          & 21.9\%          \\
                                                            & $D=4$   & 18.1\%          & 26.5\%          & 38.4\%          \\
                                                            & $D=128$ & \textbf{31.3\%} & \textbf{43.6\%} & \textbf{58.8\%} \\
        \hline
        \multirow{3}{*}{Algorithm \ref{alg:friend_sug_com}} & $D=2$   & 8.4\%           & 14.1\%          & 21.8\%          \\
                                                            & $D=4$   & 18.0\%          & 26.4\%          & 38.2\%          \\
                                                            & $D=128$ & \textbf{24.1\%} & \textbf{32.9\%} & \textbf{41.8\%} \\
        \hline
    \end{tabular}
\end{figure}

>>>>

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts

\usepackage{cite}
\usepackage[numbers]{natbib}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{xcolor}
\usepackage{graphicx}
\usepackage[warn]{textcomp}
\usepackage{multirow}
\usepackage{mathtools}
\usepackage{commath}
\usepackage{caption}
\usepackage{subcaption}
\usepackage[super]{nth}
\usepackage{url}

@@

\title{Community Embeddings for Facebook Friend Suggestions\\
    \thanks{The reported study was partially supported by RFBR grant \textnumero 20-07-00958. The paper was prepared within the framework of the HSE University Project Group Competition 2020-2022.}
}

@@

\begin{figure}[htbp]
    \centering
    \includegraphics[width=0.48\textwidth]{images/fb/graph_d4.png}
    \caption{\textit{Social circles: Facebook} graph plotted with networkx's spring layout and node colors representing community membership obtained by ComE BGMM+VI with $D=4$ \cite{networkx}.}
    \label{fig:fb_graph}
\end{figure}

@@

The function $ComE(G, K, \gamma, \ell, \zeta, D, m, \alpha, \beta)$ is the ComE BGMM+VI community embedding algorithm. It returns the node embedding $\Phi$, context embedding $\Phi'$, community assignment $\Pi$, and community embedding $(\Psi,\Sigma)$. ComE BGMM+VI is based on the ComE algorithm presented by \citeauthor{ComE} in their paper \textit{Learning Community Embedding with Community Detection and Node Embedding on Graphs} as algorithm 1 on page 381 of the publication's in-proceeding \cite{ComE}. ComE BGMM+VI takes a Bayesian approach to community embedding by utilizing  a Bayesian Gaussian mixture model (BGMM) and variational inference (VI) for community embedding and detection instead of a non-Bayesian Gaussian mixture model (GMM) with expectation maximization (EM) \cite{ComE_BGMM_GH}. %TODO citation to the ComE BGMM paper.

@@

The small \textit{Zachary’s Karate Club graph} dataset will be used to visually underline the motivation for using node embedding and community memberships generated by ComE BGMM+VI for generating friend suggestions for person nodes on the karate club graph \cite{KarateClubDS}.

\begin{figure}[htbp]
    \centering
    \includegraphics[width=0.4\textwidth]{images/karateclub/spring.png}
    \caption{\textit{Zachary’s Karate Club graph} plotted with networkx's spring layout \cite{KarateClubDS, networkx}.}
    \label{fig:karateclub_spring}
\end{figure}

Fig. \ref{fig:karateclub_spring} shows the 34 nodes and 78 edges of the \textit{Zachary’s Karate Club graph} dataset plotted as a graph with the Python network package networkx's spring layout.\cite{networkx} Spring layout simulates a force-directed representation of the network, by treating edges as springs pulling nodes together and treating nodes as repelling objects, sometimes called an anti-gravity force.\cite{networkx} The node classification, represented by node color, shown in Fig. \ref{fig:karateclub_spring} was obtained by running ComE with BGMM and VI with the hyperparameters presented in Table \ref{tab:params_karateclub}.

@@

As can be seen in Fig. \ref{fig:karateclub_spring}, the two communities in the \textit{Zachary’s Karate Club graph} dataset are correctly identified, despite the number of communities being initialized with $K=5$. Thanks to the Bayesian approach to community modeling and optimization through variational inference, three unused communities are dropped, which leaves the prominent two communities present in the dataset. The node classification obtained reflects the node classification published in the \citeyear{ComE} original ComE paper by \citeauthor{ComE}.\cite{ComE}

@@

\textit{Zachary’s Karate Club graph} is split into training and testing parts. For evaluation, the leave-one-out (LOO) method is used, where for each testing user one edge is left out, then if the left out edge is in the user's top-$N$ recommendations, a hit is counted, otherwise, a miss is recorded.

@@

\begin{figure}[htbp]
    \centering
    \includegraphics[width=1\linewidth]{images/karateclub/embeddings_5.png}
    \caption{\textit{Zachary’s Karate Club graph} training dataset node and community embedding and node classification.}
    \label{fig:karateclub_emb}
\end{figure}

@@

\begin{figure}[htbp]
    \centering
    \includegraphics[width=.4\textwidth]{images/karateclub/hit_rates.png}
    \caption{Hit-rates from top-$N$ recommendations on \textit{Zachary’s Karate Club graph}.}
    \label{fig:karateclub_hitrates}
\end{figure}

Fig. \ref{fig:karateclub_hitrates} shows promising results for using ComE to generate top-$N$ friend suggestions on the small \textit{Zachary’s Karate Club graph}. The achieved hit-rate starts at 33\% for $N=1$, reaches 67\% at $N=3$, and maxes out at 100\% for $N=7$; therefore, when presenting three friend suggestions to one test user, at least in $66\%$ of cases a friend suggestion would be of interest.

@@

The recommender system based on community embeddings with the Bayesian Gaussian mixture model and variational inference is quantitatively evaluated. Friend suggestions are generated for the \textit{Social circles: Facebook} dataset \cite{social_circles_fb, social_circles_in_ego_networks}.

The method of generating friend suggestions using ComE BGMM+VI embeddings and the smallest-distance approach is applied to the \textit{Social circles: Facebook} graph. Both Algorithm \ref{alg:friend_sug} and Algorithm \ref{alg:friend_sug_com} are evaluated. The top-$N$ recommendations are evaluated by the hit-rate metric.

@@

The \textit{Social circles: Facebook} dataset was sourced from the profile and network data of ten ego-networks (egonets) through a custom Facebook app by \citeauthor{social_circles_in_ego_networks} as part of their \citeyear{social_circles_in_ego_networks} article \textit{Learning to Discover Social Circles in Ego Networks}.\cite{social_circles_in_ego_networks}

@@

The \textit{Social circles: Facebook} graph is split into training and testing parts. For evaluation, the leave-one-out (LOO) method is used, where for each testing user one edge is left out, then if the left out edge is in the user's top-$N$ recommendations, a hit is counted, otherwise, a miss is recorded.
Users with only one connection are removed from the dataset since no network information is added by them.

$1169$ of the 4039 users of the \textit{Social circles: Facebook} graph have at least $50$ connections and are used for testing friend suggestions. The threshold of $50$ friends for considering users for testing is set to make sure enough data is available to make sensable suggestions for testing users.

@@

Community embeddings for the \textit{Social circles: Facebook} graph are generated by running ComE BGMM+VI with the hyperparameters listed in Table \ref{tab:params_fb}.

\begin{table}
    \centering
    \caption{Hyperparameters used for the \textit{Social circles: Facebook} dataset.}
    \label{tab:params_fb}
    \begin{tabular}{ l | c | r }
        parameter                    & notation & value       \\
        \hline
        \hline
        number\_walks                & $\gamma$ & $10$        \\
        \hline
        walk\_length                 & $\ell$   & $80$        \\
        \hline
        representation\_size         & $D$      & $2, 4, 128$ \\
        \hline
        num\_workers                 & \empty   & $10$        \\
        \hline
        num\_iter                    & \empty   & $3$         \\
        \hline
        com\_n\_init                 & \empty   & $10$        \\
        \hline
        reg\_covar                   & \empty   & $0.00001$   \\
        \hline
        batch\_size                  & \empty   & $50$        \\
        \hline
        window\_size                 & $\zeta$  & $10$        \\
        \hline
        negative                     & $m$      & $5$         \\
        \hline
        lr                           & \empty   & $0.025$     \\
        \hline
        alpha                        & $\alpha$ & $0.1$       \\
        \hline
        beta                         & $\beta$  & $0.1$       \\
        \hline
        down\_sampling               & \empty   & $0.0$       \\
        \hline
        communities                  & $K$      & $10$        \\
        \hline
        weight\_concentration\_prior & $\Gamma$ & $10^{-5}$   \\
    \end{tabular}
\end{table}

As is shown in Table \ref{tab:params_fb}, the embedding's representation size $D$ is varied along three values. A representation size $D \in [2,4,128]$ is chosen to determine the effect of the embedding's dimensionality on friend suggestions. It is hypothesized, that two dimensions are not sufficient to capture the structure of the \textit{Social circles: Facebook} graph, due to its size, therefore higher dimensional embeddings are evaluated as well.

@@

Fig. \ref{fig:fb_emb_anim_d2} shows node and community embeddings with representation size $D=2$. The limiting two-dimensional space does not allow ComE to represent the \textit{Social circles: Facebook} graph's structure sufficiently to utilize the $K=10$ components of the Gaussian mixture model. Only three components are used after the last of three iterations.

@@

    \caption{Hit-rates of \textit{Social circles: Facebook} graph's top-$N$ friend suggestions on $N$-axis for $D \in [2,4,128]$. Hit-rate lines for Algorithm \ref{alg:friend_sug} and Algorithm \ref{alg:friend_sug_com} overlap for $D=2$ and $D=4$.}
    \label{fig:fb_hit_rates}
\end{figure}

@@

\begin{figure}[htbp]
    \centering
    \caption{Hit-rates of \textit{Social circles: Facebook} graph's top-$N$ friend suggestions for algorithms \ref{alg:friend_sug} and \ref{alg:friend_sug_com}, $N \in [5, 10, 20]$, and $D \in [2, 4, 128]$.}
    \label{tab:fb_hit_rates}
    \begin{tabular}{ l | l | c | c | c }
        \hline
        \empty                                              & \empty  & top-$5$         & top-$10$        & top-$20$        \\
        \hline
        \hline
        \multirow{3}{*}{Algorithm \ref{alg:friend_sug}}     & $D=2$   & 8.5\%           & 14.2\%          & 21.9\%          \\
                                                            & $D=4$   & 18.1\%          & 26.5\%          & 38.4\%          \\
                                                            & $D=128$ & \textbf{31.3\%} & \textbf{43.6\%} & \textbf{58.8\%} \\
        \hline
        \multirow{3}{*}{Algorithm \ref{alg:friend_sug_com}} & $D=2$   & 8.4\%           & 14.1\%          & 21.8\%          \\
                                                            & $D=4$   & 18.0\%          & 26.4\%          & 38.2\%          \\
                                                            & $D=128$ & \textbf{24.1\%} & \textbf{32.9\%} & \textbf{41.8\%} \\
        \hline
    \end{tabular}
\end{figure}
